---
title: "warmup04-balaji-veeramani.Rmd"
author: "Balaji Veeramani"
date: "February 24, 2019"
output: html_document
---

```{r setup}
library(readr)
library(dplyr)
library(ggplot2)
```

# 1) Importing Data
```{r}
url = "https://raw.githubusercontent.com/ucb-stat133/stat133-hws/master/data/nba2018-players.csv"
dat = read_csv(url, col_types = "ccfiiiicdiiiiii")
summary(dat)
class(dat)
```

# 2) Technical Questions about "readr"
a) Here are some differences between `tibble` and `data.frame`: `tibble` prints a more concise representation of the tabe; `tibble` has more strict subsetting rules; `tibble` and `data.frame` have different subsetting rules; `tibble` never changes input types.

b) Yes. There is a column data type `col_skip` that can be used to ignore a column while reading a file. Hence, using the argument "cc------d--i---" for`col_types` would achieve the desired result.

c) The equivalent arguments for `header`, `col.names`, `na.strings`, and `colClasses` are `col_names`, `col_names`, `na`, and `col_types` respectively. Note that `col_names` can be used to both indicate whether the first line should be used as a header and the names of columns.

# 3) Salaries by Team
```{r}
team_salaries = dat %>% group_by(team) %>% mutate(salary = salary / 1000000) %>% summarise(total_salary = sum(salary), mean_salary = mean(salary), median_salary = median(salary)) %>% arrange(desc(total_salary)) %>% select(team, total_salary, mean_salary, median_salary)
as.data.frame(team_salaries)
```

```{r}
ggplot(team_salaries, aes(x = reorder(team, total_salary), y = total_salary)) + coord_flip() + geom_bar(stat = "identity") + xlab("Team") + ylab("Total Salary (in millions)") + ggtitle("NBA Teams Ranked by Total Salary") + geom_hline(yintercept = mean(team_salaries$total_salary), color = "red", size = 1)
```

# 4) Points by Team
```{r}
team_points = dat %>% group_by(team) %>% summarise(total_points = sum(points), mean_points = mean(points), median_points = median(points)) %>% select(team, total_points, mean_points, median_points) %>% arrange(desc(total_points))
as.data.frame(team_points)
```

```{r}
ggplot(team_points, aes(x = reorder(team, total_points), y = total_points)) + coord_flip() + geom_bar(stat = "identity") + xlab("Team") + ylab("Total Points") + ggtitle("NBA Teams Ranked by Total Points") + geom_hline(yintercept = mean(team_points$total_points), color = "red", size = 1)
```

# 5) Cost of Scored Points
```{r}
points_salary = inner_join(team_salaries, team_points)
summary(points_salary)
points_salary = points_salary %>% mutate(cost_point = total_salary * 1000000 / total_points)
summary(points_salary %>% select(cost_point))
```

```{r}
ggplot(points_salary, aes(x = reorder(team, cost_point), y = cost_point)) + coord_flip() + geom_bar(stat = "identity") + xlab("Team") + ylab("Cost per Point") + ggtitle("NBA Teams Ranked by Total Salary to Points Ratio") + geom_hline(yintercept = mean(points_salary$cost_point), color = "red", size = 1)
```

```{r}
ggplot(points_salary, aes(x = total_salary, y = total_points)) + geom_point(aes(color = cost_point)) + xlab("Total Points") + ylab("Total Salaries") + ggtitle("Total Salary by Total Points for NBA Teams") + geom_text(aes(label=team), vjust=-0.6, size = 2) + labs(color='Cost per Point')
```