---
title: "Introduction to binomial"
author: Balaji Veeramani
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to binomial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(binomial)
```

## Scope

The `"binomial"` package is intended for R programmers with an introductory-
level knowledge of probability concepts.

## Binomial Distributions
Binomial random variables are important across many disciplines. The binomial 
distribution is characterized by two shape parameters: `trials`, which represents 
the number of I.I.D Bernoulli trials performed; and `prob`, the probability of 
success on each I.I.D Bernoulli trial.

## Creating a `binvar`

To create a binomial variable object, use the `bin_variable` constructor:

```{r}
bernoulli <- bin_variable(1, 0.5)
bernoulli
#> "Binomial Variable" 
#>
#> Parameters 
#> - number of trials: 1 
#> - prob of success : 0.5

summary(bernoulli)
#> "Binomial Variable" 
#>
#> Parameters 
#> - number of trials: 1 
#> - prob of success : 0.5 
#>
#> Measures 
#> - mean    : 0.5 
#> - variance: 0.25 
#> - mode    : 0 
#> - skewness: 0 
#> - kurtosis: -2
```

The `bin_variable` constructor takes two arguments: `trials` and `prob`. The 
returned object is a list that contains two named vectors: `"trials"` and `"prob"`.

## Inspecting probabilities

You can calculate the probability of a single value by using the `bin_probability`
function:

```{r}
prob_heads = bin_probability(1, bernoulli$trials, bernoulli$prob)
prob_heads
#> [1] 0.5
```

To inspect the probability mass function, use the `bin_distribution` function:

```{r}
pmf = bin_distribution(bernoulli$trials, bernoulli$prob)
pmf
#>   success probability
#> 1       0         0.5
#> 2       1         0.5
```

To inspect the cumulative distribution function, use the `bin_cumulative` function:

```{r}
cdf = bin_cumulative(bernoulli$trials, bernoulli$prob)
cdf
#>   success probability cumulative
#> 1       0         0.5        0.5
#> 2       1         0.5        1.0
```

## Plots of probability distributions

The `bin_distribution` function returns a `bindis` object, and the `bin_cumulative` 
function returns a `cumdis` function. Either class of object can be plotted with 
the `plot` method.
