---
title: "warmup03-balaji-veeramani.Rmd"
author: "Balaji Veeramani"
date: "February 17, 2019"
output: html_document
---

# Importing Data
```{r}
# There is a header. Observations are separated by a comma. Missing values are codified by an empty string.
github = "https://raw.githubusercontent.com/ucb-stat133/stat133-hws/"
repo = "master/data/nba2018-players.csv"
datafile = paste0(github, repo)
data_types = c("character", "factor", "factor", rep("integer", 4), "character", "numeric", rep("integer", 6))
nba = read.csv(datafile, colClasses = data_types)
str(nba, vec.len = 1)
nba2 = read.table(datafile, header = TRUE, colClasses = data_types, sep = ",")
str(nba2, vec.len = 1)
identical(nba, nba2)
```

# Technical Questions about Importing Data
a) Each column name is the "V" character followed by the corresponding column number.
b) All column data types are factors. 
c) The default arguments for `read.table` assumes that the table is separated by whitespace. However, the NBA data is separated by commas. Hence, an errors is raised while scanning the text.
d) The option with custom-specified data types.
e) All data values are converted to strings. 

#Examine Salary
```{r}
summary(nba$salary)
hist(nba$salary)
nba$salary2 = nba$salary / 1000000
hist(nba$salary2)
nba$salary2 = log(nba$salary2)
hist(nba$salary2)
# The log-transformed distribution is more symetrical.
# The shape of the log-transformed salary2 distribution is approximately unimodal and left skewed with some outliers.
```

```{r}
low = nba$salary2 < -3
length(nba$player[low])
nba[low, c("player", "weight", "height", "team", "position")]
nba = nba[!low,]
dim(nba)
```

#Explore Points
```{r}
summary(nba$points)
hist(nba$points)
nba$points[nba$points == 0] = 0.1
hist(log(nba$points))
# The distribution of the log-transformed points is left-skewed and approximately unimodal with some outliers.
```


#Transforming Age into a Factor
```{r}
boxplot(salary2 ~ age, data = nba, xlab = "Age", ylab = "Salary (in millions)")
# The above distribution appears to be roughly unimodal and symetrical. However, an analyst could also intepret the data as roughly uniform.
boxplot(points ~ age, data = nba, xlab = "Age", ylab = "Points")
# The above distribution appears to be aproximately uniform.
```

```{r}
age2 = nba$age
age2[nba$age <= 19] = 20
age2 = cut(age2, labels = c("20-24", "25-29", "30+"), breaks = c(19, 24, 29, 100))
summary(age2)
```

#Plotting Salary against Points, by Age Group
```{r}
palette1 = c("#D4D62A", "#4F9D66", "#9575AB")
age_colors = palette1[age2]
plot(salary ~ points, data = nba, log = "xy", main = "", xlab = "Points (log scale)", ylab = "Salary (log $)", col = age_colors, pch = 19, cex = 0.6)
legend("bottomright", fill = palette1, legend = levels(age2), title = "Age Groups", cex = 0.8)
```

#Plotting Salary against Points, by Position
```{r}
boxplot(salary2 ~ position, data = nba, xlab = "Position", ylab = "Salary (in millions)")
# The salaries by position appear to be approximately uniform. In other words, there is a small difference in salary depending on position.
boxplot(points ~ position, data = nba, xlab = "Position", ylab = "Points Scored")
# The number of points scored by position appear to be approximately uniform. Point gaurds and shooting gaurds may score slightly more than the other positions.
```

```{r}
palette2 = c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3", "#a6d854")
pos_colors = palette2[nba$position]
plot(salary ~ points, data = nba, log = "xy", main = "", xlab = "Points Scored (log scale)", ylab = "Salary (log $)", col = pos_colors, pch = 19, cex = 0.6)
legend("bottomright", fill = palette2, legend = levels(nba$position), title = "Positions", cex = 0.8)
```
