---
title: "warmup05-balaji-veeramani"
author: "Balaji Veeramani"
date: "March 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(ggplot2)
```

###1) Gaussian Function
```{r}
#' @title Normal PDF
#' @description Evaluates the probability density of a normal distribution with the specified mean and standard deviation at x
#' @param x the value of the normal random variable
#' @param m the mean of the distribution
#' @param s the standard deviation of the distribution
#' @return the probability density at x
gaussian = function(x = 0, m = 0, s = 0) {
  if (s < 0) { stop("s must be greater than 0") }
  return((1 / (s * sqrt(2 * pi))) * exp(-(1/2) * ((x - m) / s)^2))
}

# test set 1
gaussian(x = 0, m = 0, s = 1)
gaussian(x = 1, m = 0, s = 2)

# test set 2
gaussian(x = 1, m = 0, s= 0)
gaussian(x = 1, m = 0, s = -1)

# gaussian curve
x_values = seq(from = -4.5, to = 4.5, by = 0.1)
y_values = gaussian(x_values, m = 0, s = 2)
plot(x_values, y_values, las = 1, type = "l", lwd = 2)
```

```{r}
# additional plot
x_values = seq(from = -5, to = 5, by = 0.01)
y_values = data.frame(x = x_values, y1 = gaussian(x_values, 0, sqrt(0.2)), y2 = gaussian(x_values, 0, sqrt(1)), y3 = gaussian(x_values, 0, sqrt(5)), y4 = gaussian(x_values, -2, sqrt(0.5)))
p = ggplot(data = y_values, aes(x)) 
p = p + geom_line(aes(y = y1, color = "#234b85"), size = 1.2) 
p = p + geom_line(aes(y = y2, color = "#ca0813"), size = 1.2) 
p = p + geom_line(aes(y = y3, color = "#c7891f"), size = 1.2)
p = p + geom_line(aes(y = y4, color = "#51991c"), size = 1.2)
p = p + labs(
  x = bquote(chi), 
  y = bquote(phi1[mu ~ "," ~ sigma^2](chi)), 
  caption = "Figure 1: Normal probability density functions") 
p = p + scale_color_manual(
  name = "", values=c("#234b85", "#51991c", "#c7891f", "#ca0813"), 
  labels = c(
    bquote(mu == 0 ~ ",   " ~ sigma^2 == 0.2 ~ ","), 
    bquote(mu == 0 ~ ",   " ~ sigma^2 ==~ "1.0,"), 
    bquote(mu == 0 ~ ",   " ~ sigma^2 == ~ "5.0,"), 
    bquote(mu == -2 ~ ", " ~ sigma^2 == 0.5 ~ ","))) 
p = p + guides(color = guide_legend(label.position = "left", keywidth = 1.7)) 
p = p + theme_minimal() 
p = p + theme(
  title = element_text(size = 16),
  plot.caption = element_text(size = 12.5, hjust = 0.12, vjust = -2.25),
  axis.text.y = element_text(margin = margin(r = 1), hjust = 3),
  legend.justification = c(1, 0), 
  legend.position = c(0.98, 0.67), 
  legend.text.align = 0,
  legend.title = element_blank(),
  legend.text = element_text(size = 10),
  legend.background = element_rect(color = "white", fill = alpha("white", 1), siz),
  panel.grid.major = element_line(), 
  panel.grid.minor = element_blank(), 
  panel.border = element_rect(colour = "black", fill=NA, size=1)) 
p = p + scale_x_continuous(breaks = pretty(y_values$x, n = 10)) 
p = p + scale_y_continuous(breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1, 1.2), limits = c(0, 1))
p
```

###2) Descriptive Statistics
```{r}
#' @title Descriptive Statistics
#' @description Returns descriptive statistics about a numeric vector
#' @param x a numeric vector
#' @param na.rm a logical value indicating whether NA values should be stripped
#' @return descriptive statistics about the specified numeric vector
descriptive = function(x = 0, na.rm = FALSE) {
  if (mode(x) != "numeric") { stop("input vector must be numeric") }
  metrics = c(
    min(x, na.rm = na.rm),
    quantile(x, na.rm = na.rm)[2],
    median(x, na.rm = na.rm),
    mean(x, na.rm = na.rm),
    quantile(x , na.rm = na.rm)[4],
    max(x, na.rm = na.rm),
    max(x, na.rm = na.rm) - min(x, na.rm = na.rm),
    quantile(x, na.rm = na.rm)[4] - quantile(x, na.rm = na.rm)[2],
    sd(x, na.rm = na.rm)
  )
  names(metrics) = c("min", "q1", "median", "mean", "q3", "max", "range", "iqr", "sd")
  return(metrics)
}

set.seed(100)
x = rnorm(100)
y = x
y[sample(1:100, size = 10)] = NA

# test set 1
descriptive(x)
descriptive(y)
descriptive(y, na.rm = TRUE)
descriptive(letters)

# test set 2
a = descriptive(x)
class(a)
length(a)
names(a)
```

###3) Minkowski Distance
```{r}
#' @title Minkowski Distance
#' @description Calculates the Minkowski Distance between two points
#' @param x a numeric vector 
#' @param y a numeric vector
#' @return the Minkowski Distance between x and y
minkowski = function(x, y, p = 1) {
  if (length(x) != length(y)) { stop("x and y have different lengths") }
  if (mode(p) == "numeric" && p < 1) { stop("p cannot be less than 1")}
  if (mode(p) == "character" && p != "max") { stop("invalid character value for p")}
  
  if (mode(p) == "numeric") {
    return(sum(abs(x - y)^p)^(1/p))
  } else{ 
    return(max(abs(x - y)))
  }
}

point1 = c(0, 0)
point2 = c(1, 1)
point3 = sqrt(c(2, 2))
point4 = c(0, 1)
point5 = c(1, 1, 1)

# test set 1
minkowski(point1, point2, p = 1)
minkowski(point1, point3, p = 2)
minkowski(point1, point2, p = "max")

# test set 2
minkowski(point4, point5, p = 1)
minkowski(point1, point2, p = 0.5)
minkowski(point1, point2, p = "min")