---
title: "warmup02-balaji-veeramani.Rmd"
author: "Balaji Veeramani"
date: "February 10, 2019"
output: html_document
---

###1) Vectors
```{r}
number = c(30L, 35L, 23L, 9L, 11L, 27L, 34L, 6L, 3L, 0L)
player = c("Stephen Curry", "Kevin Durant", "Draymond Green", "Andre Iguodala", "Klay Thompson", "Zaza Pachulia", "Shaun Livingston", "Nick Young", "David West", "Patrick McCaw")
position = factor(c("PG", "PF", "PF", "SF", "SG", "C", "PG", "SG", "C", "SG"))
height = c("6-3", "6-9", "6-7", "6-6", "6-7", "6-11", "6-7", "6-7", "6-9", "6-7")
weight = c(190, 240, 230, 215, 215, 270, 192, 210, 250, 185)
birthdate = c("March 14, 1988", "September 29, 1988", "March 4, 1990", "January 28, 1984", "February 8, 1990", "February 10, 1984", "September 11, 1985", "June 1, 1985", "August 29, 1980", "October 25, 1995")
experience = c(8L, 10L, 5L, 13L, 6L, 14L, 12L, 10L, 14L, 1L)
college = c("Davidson College", "University of Texas at Austin", "Michigan State University", "University of Arizona", "Washington State University", NA, NA, "University of SOuthern California", "Xavier University", "University of Nevada, Las Vegas")
```

```{r}
player[which.max(weight)]
college[height == "6-6"]
position[which.max(experience)]
number[which.min(weight)]
table(position)
length(player[weight > mean(weight)])
length(player[experience >= 9 & experience <= 12])
mean(experience[position == "SG"])
median(weight[position != "C"])
quantile(experience[position == "PF" | position == "SG"])[2]
```

###2) Lists for GSW
```{r}
gsw = list(
  player = player,
  number = number,
  position = position,
  weight = weight,
  experience = experience
)
```

```{r}
gsw$number[which.max(gsw$weight)]
gsw$position[which.min(gsw$experience)]
length(gsw$player[gsw$experience < 8 | gsw$experience > 11])
quantile(gsw$experience[gsw$position == "PF" | gsw$position == "SG"])[4]
weight_deviation = abs(gsw$weight - mean(gsw$weight))
gsw$player[which.max(weight_deviation)]
```

###3) More lists
```{r}
hp = list(
  first = "Harry",
  last = "Potter",
  courses = c("Potions", "Enchantments", "Spells"),
  sport = "Quidditch",
  age = 18L,
  gpa = 3.9
)
```

```{r}
class(hp)
length(hp)
length(hp$courses)
typeof(hp$age)
typeof(hp$gpa)
typeof(c(hp$age, hp$gpa))
```

###4) Technical Questions
*a)* The R coercion rules state that numeric types are ordered higher than logical types. The `TRUE` value is of logical type. Therefore, the `TRUE` value is coerced to a double. When the `TRUE` boolean  is coerced to a double, the value is implicitly converted into the integer 1. Consequently, the expression evaluates to 2.

*b)* The R coercion rules state that characters are ordered higher than numeric types. The 0 literal is a double. Therefore, the 0 object is coerced to a character. By the lexigraphic collation ordering, the character "-" is ordered before "0". As a result, the expression evaluates to `FALSE`.

*c)* The R compiler determines that the expression `(10 <= 5)` is evaluated first. Clearly this expression evaluates to `FALSE`. We then have the expression `FALSE >= 0`. `FALSE` is coerced to 0, which yields the expression `0 >= 0`. Hence, the original expression evaluates to `TRUE`.

*d)* In R, the unary negation operator is of lower precedence that the binary addition operator. Therefore, in the first command the add function is applied to 1 and the negation of `TRUE`, whereas in the second command the negation operator is applied to the result of adding `TRUE` and 1. The results of each expression is therefore different.

###5) Subsetting
```{r}
lord = c("v", "o", "l", "d", "e", "m", "o", "r", "t")
```

*a)* The `TRUE` valued is recyled to match the length of the `lord` vector. Hence, the vector `TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE` is formed. By the subsetting rules, every element of `lord` is return. Hence a copy of the `lord` vector is produced.
``` {r}
lord[TRUE]
```

*b)* The expression evaluates to the logical constant `NA` because the subset index is greater than the length of the vector. 
``` {r}
lord[length(lord) + 1]
```

*c)* The `seq` function with the specified arguments returns a numeric vector. The first element of the vector is nine, which is the length of the `lord` vector, and subsequently includes every other number in the range 9, 8, ..., 1. When the `lord` vector is subsetted with the previosuly described vector, every other element of `lord` in reverse is produced.
``` {r}
lord[seq(from = length(lord), to = 1, by = -2)]
```

*d)* The expression `lord == "o"` evaluates to the logical vector `FALSE TRUE FALSE FALSE FALSE FALSE TRUE FALSE FALSE`. Subsetting the `lord` vector with the logical vector then produces the vector `"o", "o"`. 
```{r}
lord[lord == "o"]
```

*e)* R first evaluates the expression `lord != "e"`, which evaluates to the logical vector `TRUE TRUE TRUE TRUE FALSE TRUE TRUE TRUE TRUE`. Next, the expression `lord != "o"` is evaluated to `TRUE FALSE TRUE TRUE TRUE TRUE FALSE TRUE TRUE`. Applying the `&` operator to these operators produces the logical vector of characters that are neither "e" nor "o": `TRUE FALSE TRUE TRUE FALSE TRUE FALSE TRUE TRUE`. Subsetting the `lord` vector using the prior logical vector produces the desired vector.
```{r}
lord[lord != "e" & lord != "o"]
```

*f)* The vectorized `%in%` operator applied to the `lord` vector and the vector of vowels produces the vector describing whether each element of `lord` in a vowel. Subsetting `lord` with the logical vector produces the vector of vowels in `lord`.
```{r}
lord[lord %in% c('a', 'e', 'i', 'o', 'u')]
```

*g)* As described in the passage above, the expression `lord %in% c('a', 'e', 'i', 'o', 'u')` evaluates to a logical vector describing whether each element of `lord` is a vowel. Applying the vectorized negation operaton on the vector produces the logical vector representing which elements of `lord` are consonants. Subsetting using this vector produces the vector of consonants in `lord`, which is subsequently converted to uppercase by the `toupper` function.
```{r}
toupper(lord[!(lord %in% c('a', 'e', 'i', 'o', 'u'))])
```

*h)* The `paste` converts its arguments to characters. When the `collapse` argument is not null, a scalar is produced with the elements of the original vector delimited by the argument. Hence, the expression `paste(lord, collapse = '')` evaluates to the string `voldemort`.
```{r}
paste(lord, collapse = '')
```

*i)* The vectorized function `is.na` produces a vector that describes if each is element is the `NA` constant. Since none of the elements of `lord` are `NA`, subsetting `lord` with the result of `is.na(lord)` produces an empty character vector, namely `character(0)`.
```{r}
lord[is.na(lord)]
```

*j)* The expression `is.na(lord)` is evaluated first. Clearly, the expression evaluates to a logical vector such that all elements are equal to `FALSE`. Applying the vectorized negation operator to vector produces a vector containing only `TRUE`. The `sum` function expects a numeric type. Hence, the logical vector of `TRUE` is coerced to a vector of ones. Since the length of `lord` is nine, the original expression evaluates to 1 * 9 = 9. 
```{r}
sum(!is.na(lord))
```

###6) 2D Arrays (i.e Matrices)
```{r}
vol = matrix(lord, nrow = 3, ncol = 3)
vol
vol[1,]
vol[1:2, 2:1]
vol[3:1, ]
vol[,c(1, 2, 2)]
vol[3:1, 3:1]
vol[c(3:1, 1:3), c(3, 2, 2, 3)]
vol[c(3:1, 1:3), c(1:3, 3:1)]
