---
output:
  github_document:
    html_preview: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## Overview

`"binomial" is an R package that provides basic tools for simulating
and analyzing binomial random variables.

* `bin_variable()` constructs an object that represents a binomial random variable.
* `bin_probability()` returns the probability that a binomial random variable with 
  the specified shape parameters take on a particular value.
* `bin_distribution()` returns the probability mass function of a binomal random 
  variable with the specified shape parameters
* `bin_cumulative()` returns the cumulative distribution function of a binomal random 
  variable with the specified shape parameters
* `summary()` method of `binvar` objects
* `plot()` method of `bindis` and `cumdis` objects

## Motivation

The purpose of the "binomial" package is to provide minimal tools for
simulative and analyzing binomial random variables.

## Installation

Install the "binomial" package from GitHub with the following commands:

```r
install.packages("devtools") 
devtools::install_github("stat133-sp19/hw-stat133-bveeramani/workout03/binomial")
```

## Usage

```{r}
library(binomial)

bernoulli <- bin_variable(1, 0.5)
bernoulli
#> "Binomial Variable" 
#>
#> Parameters 
#> - number of trials: 1 
#> - prob of success : 0.5

bin_distribution(bernoulli$trials, bernoulli$prob)
#>   success probability
#> 1       0         0.5
#> 2       1         0.5

bin_probability(1, bernoulli$trials, bernoulli$prob)
#> [1] 0.5

summary(bernoulli)
#> "Binomial Variable" 
#>
#> Parameters 
#> - number of trials: 1 
#> - prob of success : 0.5 
#>
#> Measures 
#> - mean    : 0.5 
#> - variance: 0.25 
#> - mode    : 0 
#> - skewness: 0 
#> - kurtosis: -2

cdf = bin_distribution(bernoulli$trials, bernoulli$prob)
plot(cdf)
```
